Estimating the number of clusters using Cross Validation

### 2015-06-16 cvclust.tex
add simulation results to main doc (needs reformatting)
start generalization of Pollard's thm

\newcommand{\CV}{\operatorname{CV}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}
\newcommand{\bpi}{\bar \pi}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}

@@

\section{Theoretical justification of Gabriel CV method}
This section gives the self-consistency proof of the proposed Gabriel method. What's more, its asymptotic property under suitable condition will be shown in this section in order to give the idea of why the Gabriel CV method works.\\

@@

@@

\begin{enumerate}[label={Assumption},leftmargin=*]
	\item $1.$ There are $K$ true centers $(\mu_1,...,\mu_K)$ where $\mathbf{X}$ generated from
	\item $2.$ The $K$ partitioned centers $(\mu^L_1,...,\mu^L_K)$ are all distinct in $L$ dimensions: $||\mu^L_i-\mu^L_j|| \neq 0$ for any $i \neq j$; $(\mu^R_1,...,\mu^R_K)$ are all distinct in $R$ dimensions
	\item $3.$ There is noise, i.e. $$x_i = \mu \left(c_i\right) + \varepsilon_i, \hspace{0.2 in} \mu \left(c_i\right) \in \{\mu_1,...,\mu_K\}$$ where vector $\varepsilon_i$ is i.i.d with zero mean and covariance matrix $\Sigma$, which is independent from any $\mu_j \in (\mu_1,...,\mu_K)$, $c_i$ is the true cluster for observation $i$, which implies $$x^L_i = \mu^L \left(c_i\right) + \varepsilon^L_i,\hspace{0.2in} x^R_i = \mu^R \left(c_i\right) + \varepsilon^R_i$$ 
with $$\mu^L\left(c_i\right) \in \{\mu^L_1,...,\mu^L_K\},\hspace{0.2in} \mu^R\left(c_i\right) \in \{\mu^R_1,...,\mu^R_K\} $$\
where $\varepsilon^L_i$ and $\varepsilon^R_i$ are independent and both have zero mean and covariance matrix $\Sigma^L$ and $\Sigma^R$ respectively, so $\Sigma = 
\begin{pmatrix}
  \Sigma^L &0\\
  0  &\Sigma^R\\
  \end{pmatrix}$
. Also $\varepsilon^R_i$ is bounded $||\varepsilon^R_i|| \leq 1$.
	\item $4.$ Clusters are well separated, i.e. \[ ||x_i-\mu \left(c_i\right)|| < ||x_i-\mu_{j}|| \hspace{0.2 in}\forall \hspace{0.01 in}j\neq c_i \] and \[ ||x^L_i-\mu^L \left(c_i\right)|| < ||x^L_i-\mu^L_{j}||-\delta,  \hspace{0.2 in}||x^R_i-\mu^R \left(c_i\right)|| < ||x^R_i-\mu^R_{j}||-\delta \hspace{0.2 in}\forall \hspace{0.01 in}j\neq c_i\] where $\delta > 0$ is a small constant 
	\item $5.$ The training set contains $n^{train}_j>0$ observations from each cluster $j$ and the test set contains $n^{test}_j > 0$ observations from each cluster $j$, $j=1,2...,K$
	\item $6.$ The prior for $x_i$ come from cluster $\mu_j$ is $\pi_j > 0$. So when the total number of observation in training set $m$ goes infinity, $\lim_{m\to \infty} n^{train}_j = m \cdot \pi_j $. Also $m=n\cdot r$, $n-m = n\cdot (1-r)$, where $0<r<1$ is a positive constant.

@@

@@

\caption{ \label{table1}Simulation results} 

>>>>

\newcommand{\CV}{\operatorname{CV}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}
\newcommand{\bpi}{\bar \pi}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}
\newcommand{\sE}{\mathcal{E}}
\newcommand{\sX}{\mathcal{X}}

@@

\section{Theoretical justification of Gabriel CV method}

This section gives the self-consistency proof of the proposed Gabriel method. What's more, its asymptotic property under suitable condition will be shown in this section in order to give the idea of why the Gabriel CV method works.\\

@@

\subsection{Single Cluster With Noise}

Here, we show that Gabriel Cross-Validation is consistent when the data come
from a single cluster; the algorithm picks $k = 1$ with probability tending to
$1$ as the number of rows in the data matrix $\dataX$ increase.

Our proof
relies on results from \citet{pollard1981strong}, who analyzed $k$-means and
showed that the within-cluster sum of squares converges to a population
quantity.  Specifically, for any finite set $A \subset \mathbb{R}^d$ and
probability measure $P$ supported on $\mathbb{R}^d$, define the population
within-cluster sum of squares as
\[
  W(A, P) = \int \min_{a \in A} \|x - a\|^2 P(dx).
\]
Set
\[
  m_k(P) = \inf\{ W(A, P) : \text{$A$ contains $k$ or fewer points} \}.
\]
Suppose that $X_1, \dotsc, X_n$ are independent draws from $P$, and
if $P_n$ is the empirical measure formed by placing mass $1/n$ on each $X_i$.
Let $A_n$ be a set minimizing $W(A_n, P_n)$.  Pollard gave conditions for
$A_n$ to converge to the minimizer of $W(\cdot, P)$.  His main theorem assumes that
the minimizer of $W(\cdot, P)$ is unique, which is too strong for our purposes
(this fails, for example, with a bivariate normal distribution when $k \geq 2$).
It is straightforward to generalize his result to cases with multiple
maximizers.  To do so, we will need the following three lemmas.  The proofs of
these lemmas are given in Sections~3 and~4 of Pollard's article.

\begin{lemma}\label{lem:pollard-compact}\label{lem:pollard-first}
Suppose that $\int \|x\|^2 P(dx) < \infty$ and that
$m_j(P) > m_{j+1}(P)$ for $j = 1, \dotsc, k-1$.  If $A_n$ minimizes
$W(\cdot, P_n)$, then, there exists a finite
$M$ such that as if $n$ is large enough, then, almost surely, $A_n$ is in
the class of sets $\sE_k = \{ A \subset B(5M) : \text{$A$ contains $k$ or fewer
points} \}$, where $B(5M)$ denotes the Euclidean ball of radius $5M$ centered
at the origin.
\end{lemma}
\begin{proof}
Follows from Pollard's argument in Section 3 of his paper.
\end{proof}

\begin{lemma}\label{lem:polalrd-uniform}
If $\int \|x\|^2 P(x) < \infty$, then
\(
  \sup_{A \in \sE_k} |W(A, P) - W(A,P_n)| \to 0\text{ a.s.}
\)
\end{lemma}
\begin{proof}
Given at the beginning of Pollard's Section~4.
\end{proof}

\begin{lemma}\label{lem:pollard-continuous}\label{lem:pollard-last}
If $\int \|x\|^2 P(x) < \infty$, then
the map $A \mapsto W(A, P)$ is continuous on $\sE_k$ with respect to the
topology induced by the Hausdoff metric $H(\cdot, \cdot)$, defined for compact
subsets $A, B$ of $\mathbb{R}^d$ by $H(A,B) < \delta$ if and only if every
point of $A$ is within Euclidean distance $\delta$ of at least one point of
$B$, and vice versa.
\end{lemma}
\begin{proof}
Given at the end of Pollard's Section~4.
\end{proof}

Now, define $\mathcal{\bar A} = \mathcal{\bar A}(k)$ to be the class of sets
$\bar A$ with $k$ or fewer points such that $W(\bar A,P) = m_k(P)$.
Lemmas~\ref{lem:pollard-compact} and~\ref{lem:pollard-continuous} together
with the extreme value theorem imply that this class is nonempty.

To prove
this, it suffices to analyze a single fold of the cross-validation procedure.
We will assume the following setup:

Observations $\{ (X_i, Y_i) \}_{i=1}^{n+N}$ are independent and identically
distributed.  The distribution for each $(X_i, Y_i)$ pair is such that $X_i$
and $Y_i$ have marginal distribution $P^X$ and $P^Y$, and the vectors
$X_i$ and $Y_i$ are independent.

@@

\begin{enumerate}[label={Assumption},leftmargin=*]
	\item $1.$ There are $K$ true centers $(\mu_1,...,\mu_K)$ where $\mathbf{X}$ generated from
	\item $2.$ The $K$ partitioned centers $(\mu^L_1,...,\mu^L_K)$ are all distinct in $L$ dimensions: $||\mu^L_i-\mu^L_j|| \neq 0$ for any $i \neq j$; $(\mu^R_1,...,\mu^R_K)$ are all distinct in $R$ dimensions
	\item $3.$ There is noise, i.e. $$x_i = \mu \left(c_i\right) + \varepsilon_i, \hspace{0.2 in} \mu \left(c_i\right) \in \{\mu_1,...,\mu_K\}$$ where vector $\varepsilon_i$ is i.i.d with zero mean and covariance matrix $\Sigma$, which is independent from any $\mu_j \in (\mu_1,...,\mu_K)$, $c_i$ is the true cluster for observation $i$, which implies $$x^L_i = \mu^L \left(c_i\right) + \varepsilon^L_i,\hspace{0.2in} x^R_i = \mu^R \left(c_i\right) + \varepsilon^R_i$$ 
with $$\mu^L\left(c_i\right) \in \{\mu^L_1,...,\mu^L_K\},\hspace{0.2in} \mu^R\left(c_i\right) \in \{\mu^R_1,...,\mu^R_K\} $$\
where $\varepsilon^L_i$ and $\varepsilon^R_i$ are independent and both have zero mean and covariance matrix $\Sigma^L$ and $\Sigma^R$ respectively, so $\Sigma = 
\begin{pmatrix}
  \Sigma^L &0\\
  0  &\Sigma^R\\
  \end{pmatrix}$
. Also $\varepsilon^R_i$ is bounded $||\varepsilon^R_i|| \leq 1$.
	\item $4.$ Clusters are well separated, i.e. \[ ||x_i-\mu \left(c_i\right)|| < ||x_i-\mu_{j}|| \hspace{0.2 in}\forall \hspace{0.01 in}j\neq c_i \] and \[ ||x^L_i-\mu^L \left(c_i\right)|| < ||x^L_i-\mu^L_{j}||-\delta,  \hspace{0.2 in}||x^R_i-\mu^R \left(c_i\right)|| < ||x^R_i-\mu^R_{j}||-\delta \hspace{0.2 in}\forall \hspace{0.01 in}j\neq c_i\] where $\delta > 0$ is a small constant 
	\item $5.$ The training set contains $n^{train}_j>0$ observations from each cluster $j$ and the test set contains $n^{test}_j > 0$ observations from each cluster $j$, $j=1,2...,K$
	\item $6.$ The prior for $x_i$ come from cluster $\mu_j$ is $\pi_j > 0$. So when the total number of observation in training set $m$ goes infinity, $\lim_{m\to \infty} n^{train}_j = m \cdot \pi_j $. Also $m=n\cdot r$, $n-m = n\cdot (1-r)$, where $0<r<1$ is a positive constant.
\end{enumerate}

@@

\begin{table}
\centering
\caption{ \label{table1}Simulation results} 
{\scriptsize
\begin{tabular}{lcccccccccccccccccc}
\hline
Number of Clusters & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & NA & PE \\ \hline
\input{demo/bench/setting1/results.tex}
\hline
\input{demo/bench/setting2/results.tex}
\hline
\input{demo/bench/setting3/results.tex}
\hline
\input{demo/bench/setting4/results.tex}
\hline
\input{demo/bench/setting5/results.tex}
\hline
\input{demo/bench/setting6/results.tex}
\hline
\end{tabular}
}\\
\hspace{-2.0in} \footnotesize {In Setting~1, ``\textendash'' means the method can not be used with parameter $k=1$}.
\end{table}

@@

\caption{ \label{table1}Old Simulation results}
